---
title: "Condition Report Metrics"
author: "Ben Best"
date: "July 28, 2016"
output: 
  html_document: 
    toc: yes
    toc_float: true
    code_folding: hide
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- source: http://sanctuaries.noaa.gov/science/condition/monterey-bay-2015/mbnms-cr-2015-highres.pdf -->

## TEST

### map

Let's do a little interactive map...

```{r map}
library(leaflet)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=-82.633365, lat=27.760501, popup="IMARS")
m  # Print the map
```

### time series

```{r time series}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```


## WATER

### 1. Are specific or multiple stressors, including changing oceanographic and atmospheric conditions, affecting water quality?

### 2. What is the eutrophic condition of sanctuary waters and how is it changing?

### 3. Do sanctuary waters pose risks to human health?

### 4. What are the levels of human activities that may influence water quality and how are they changing?

## HABITAT

### 5. What is the abundance and distribution of major habitat types and how is it changing?

### 6. What is the condition of biologically-structured habitats and how is it changing?

### 7. What are the contaminant con- centrations in sanctuary habitats and how are they changing?

### 8. What are the levels of human activities that may in uence habitat quality and how are they changing?

## LIVING RESOURCES

### 9. What is the status of biodiversity and how is it changing?

### 10. What is the status of non-indigenous species and how is it changing?

### 11. What is the status of key species and how is it changing?

### 12. What is the condition or health of key species and how is it changing?

### 13. What are the levels of human activities that may in uence living resource quality and how are they changing?

## MARITIME ARCHAEOLOGICAL RESOURCES

### 14. What is the integrity of known maritime archaeological re- sources and how is it changing?

### 15. Do known maritime archaeological resources pose an environmental hazard and is this threat changing?

### 16. What are the levels of human activities that may in uence maritime archaeological resource quality and how are they changing?

